### 2013-02-14 doc/icml2013/algo.tex

\section{A Method of Moments Algorithm}
\label{sec:algo}

In this section, we will outline an algorithm to recover the parameters
$\pi$ and $B$ of the mixture of linear regressions model.

% Motivate how we can recover $B$ from regression by lower order moments
Our goal is to recover the parameters $(\pi, B)$, ideally from lower
order moments. Conventionally, when using the method of moments, the
moments of observed variables are expressed algebraically in terms of
the model parameters. Using these equations, we attempt to solve for the
model parameters using empirical estimates of the moments.

Let us consider the response variable $y$, 
\begin{align}
  y &=& \innerp{\beta_h}{x} + \epsilon,
\end{align}
where $\epsilon \sim \normal{0}{\sigma^2}$ is the observation noise.  It
is natural to consider regression on the data $\{\xni,
\yni\}_{i=1}^{N}$, but it would only help us find the mean regression
coefficients, $M_1 = \E[\beta_h] = \sum_h \pi_h \beta_h$, which are insufficient to
identify the model. Let us consider some more low powers of $y$, noting
that taking the expectation gives us the corresponding low-order
moments,
\begin{align}
  y^2 &= \innerp{\beta_h\tp{2}}{x\tp{2}} + \epsilon \innerp{\beta_h\tp{2}}{x\tp{2}} + \epsilon^2 \label{eq:y2} \\ 
  y^3 &= \innerp{\beta_h\tp{3}}{x\tp{3}} + 3 \epsilon \innerp{\beta_h\tp{2}}{x\tp{2}} \notag \\
  &+ 3 \epsilon^2 \innerp{\beta_h}{x} + \epsilon^3. \label{eq:y3} 
\end{align}
Regression with the second order moments would give us $M_2 = \E[\beta_h
\tp{2}] = \sum_h \pi_h \beta_h\tp{2}$ which can identify $\pi$ and the
subspace spanned the $\beta_h$, but not the $\beta_h$ themselves. To see
this, note that the eigen-decomposition of $M_2$ is not unique.
\todo{explain better}.

@@

>>>>

\section{Spectral Experts Algorithm}
\label{sec:algo}

In this section, we outline our Spectral Experts algorithm
for recover the model parameters $\theta = (\pi, B, \sigma^2)$.
Our algorithm consists of two parts:
(i) low-rank regression to estimate certain symmetric tensors;
and (ii) tensor factorzation to recover the parameters.
% Motivate how we can recover $B$ from regression by lower order moments
%ideally from lower order moments.
%Conventionally, when using the method of moments, the
%moments of observed variables are expressed algebraically in terms of
%the model parameters. Using these equations, we attempt to solve for the
%model parameters using empirical estimates of the moments.

%%% first moment
To warm-up, let us consider performing linear regression
on the response $y$ given $x$:
\begin{align}
  y &= \innerp{\bar \beta}{x} + (\innerp{\bar\beta - \beta_h}{x} + \epsilon),
\end{align}
where $\bar\beta \eqdef \sum_{h=1}^K \pi_h \beta_h$ is the average regression coefficient.
Here, there are two noise terms:
(i) the \emph{mixing noise} $\innerp{\bar\beta - \beta_h}{x}$
due to the choice of the mixture component $h$
and (ii) the \emph{observation noise} $\epsilon \sim \normal{0}{\sigma^2}$.
Performing regression on the data $\{\xni,
\yni\}_{i=1}^{N}$ would provide a consistent estimate of $\bar\beta$,
but this is clearly insufficient for identifying all the parameters $\theta$.

%%% second moment
Regressing on $x$ provides only first-order
information.  The key insight is that to get more information about the
parameters, we need to consider regressing on higher order powers,
analogous to consider higher order moments in method of moments estimators.
Let us then consider regressing $y^2$ on $x^\otimes$:
\begin{align}
  y^2 &= \innerp{\beta_h\tp{2}}{x\tp{2}} + \epsilon \innerp{\beta_h\tp{2}}{x\tp{2}} + \epsilon^2 \label{eq:y2}.
\end{align}

Let us consider some more low powers of $y$, noting
that taking the expectation gives us the corresponding low-order
moments,
\begin{align}
  y^2 &= \innerp{\beta_h\tp{2}}{x\tp{2}} + \epsilon \innerp{\beta_h\tp{2}}{x\tp{2}} + \epsilon^2 \label{eq:y2} \\ 
  y^3 &= \innerp{\beta_h\tp{3}}{x\tp{3}} + 3 \epsilon \innerp{\beta_h\tp{2}}{x\tp{2}} \notag \\
  &+ 3 \epsilon^2 \innerp{\beta_h}{x} + \epsilon^3. \label{eq:y3} 
\end{align}
Regression with the second order moments would give us $M_2 = \E[\beta_h
\tp{2}] = \sum_h \pi_h \beta_h\tp{2}$ which can identify $\pi$ and the
subspace spanned the $\beta_h$, but not the $\beta_h$ themselves. To see
this, note that the eigen-decomposition of $M_2$ is not unique.
\todo{explain better}.

@@

\cite{candesPhaseLift} regresses on the square of the response,
because that's the observation.

### 2013-02-14 doc/icml2013/intro.tex

% Latent variables are important
% Local optima is problem
% Recent method of moments spectral methods sidestep local optima

Latent variable models have routinely been used in probabilistic
modelling, (examples). Parameters for these models have been learnt
through a variety of methods, EM, variational methods, etc. The local
nature of these algorithms lead them to converge to local optima.
Recently, there has been an interest in method of moments estimators for
generative models such as the hidden Markov
model\cite{AnandkumarHsuKakade2012} or probabilistic context free
grammars\cite{HsuKakadeLiang2012}.These estimators feature polynomial
sample and computational complexities while being statistically
consistent, i.e. given sufficient data they are guaranteed to converge
to the global optimum in an identifiable model, thus sidestepping the
local optima issue altogether.

% Discriminative models work better in practice Discriminative models
% with latent variables add expressive power (examples)
In practice it has been observed that discriminative models work better
in terms of parameter recovery\citationneeded, particularly in the large
sample regime. \todo{Examples! CRFs, etc.}
% This paper: first step in establishing provably correct estimation
% with simple mixture of linear regression
In this work, we provide a first step in establishing provably correct
estimation for such discriminative models by proposing an algorithm for
the simple mixture of linear regressions. We describe the model formally
in more detail in \sectionref{sec:model}.

@@

% Approach: Derive moment functions based on linear regression on Y^2
% and Y^3 Algorithm: trace norm regularization to recover tensor
% + tensor factorization
We propose an algorithm in \sectionref{sec:algo} that derives moment
functions based on linear regression on the conditional moments of $y$,
i.e. $\E[y^2 | X]$ and $\E[y^3 | X]$. We exploit the low rank properties
of the second and third order moments to efficiently recover parameters,
and then recover the individual regressors, $\beta_k$, using the tensor
factorization method \cite{AnandkumarGeHsu2012}.

% Show some identifiability results; require quadratic and cubic
% independence
We prove that this approach is indeed statistically consistent with
polynomial sample complexity in \sectionref{sec:theory}. Aside from
standard identifiability results \citationneeded, we also show that
recovery using moments places some independence requirements on the
variables $X$.

@@

We evaluate our algorithm on simulated linear and non-linear data to
understand how well the algorithm scales with the number of components,
dimensions and separation. To show how robust the algorithm is to model
misspecification, we evaluate it's performance on \todo{a motion
tracking dataset(?)}.

@@

We compare our approach with expectation-maximization; our results,
presented in \sectionref{sec:evaluation} shows that spectral methods do
indeed provide an answer close enough to the global optima that using
a local method from there recovers the true parameters. \todo{Describe
some of our other high-level findings.}

\todo{What about talking about initialization? Generalizing and kmeans++}

@@

We will use the notation $x\tp{p}$ to describe the $p$-th order tensor
formed by taking the outer product of $x$; i.e. $x\tp{p}_{i_1 i_2 \ldots
i_p} = x_{i_1} x_{i_2} \cdots x_{i_p}$. We will use
$\innerp{\cdot}{\cdot}$ to denote the general dot product between $p$-th
order tensors, i.e. $\innerp{X}{Y} = \sum_{i_1, i_2, \ldots i_p} X_{i_1,
i_2, \ldots i_p} Y_{i_1, i_2, \ldots i_p}.$ Similar to how $x^T A$ is
the projection of a vector $x$ onto the row space of a matrix $A$, when
$X \in \Re^{d^p}$ and $Y \in \Re^{d^{p+q}}$,  with $p, q > 0$,
$\innerp{X}{Y} \in \Re^{d^q}$ is the projection of $X$ onto the first
$p$ dimensions of $Y$, i.e. $(\innerp{X}{Y})_{i_1, \ldots, i_q}
= \sum_{j_1, \ldots, j_p} X_{j_1, \ldots, j_p} Y_{j_1, \ldots, j_p, i_1,
\ldots, i_q}$.

The expected moments of the parameters $\beta_1, \dots, \beta_k$ play an
important role in our algorithm and we will use $M_p = \E[\beta\tp{p}]
= \sum_{k=1}^{K} \pi_k \beta_k^{\otimes p}$ to denote them.

>>>>

% Discriminative models work better in practice Discriminative models
% with latent variables add expressive power (examples)
Discriminative latent-variable models,
which combine the high accuracy of discriminative models
with the compact expressiveness of latent-variable models,
have been widely applied to many tasks, including
object recognition \cite{quattoni04nips},
human action recognition \cite{wang09},
syntactic parsing \cite{petrov08discriminative},
and machine translation \cite{liang06discrimative}.
% http://www.cs.sfu.ca/~mori/research/papers/wang_cvpr09.pdf
%Slav Petrov, Discriminative Log-Linear Grammars with Latent Variables
However, parameter estimation in these models is difficult;
past approaches have relied exclusively on local optimization (EM, 
gradient descent) and are vulerable to local optima.
%Our goal is to develop consistent estimators which sidestep

% Goal of this paper
% This paper: first step in establishing provably correct estimation
% with simple mixture of linear regression
Our broad goal is to develop efficient provably consistent estimators for
discriminative latent-variable models.
In this paper, we provide a first step in this 
direction by proposing a new algorithm for a simple model,
\emph{mixture of linear regressions} (MLG) \cite{VieleTong2002}.
%mixture of linear regressions (\sectionref{sec:model}).

% Recent method of moments spectral methods sidestep local optima
Recently, method of moments estimators have been developed for
\emph{generative} latent-variable models, including
mixture models, HMMs \cite{AnandkumarHsuKakade2012},
Latent Dirichlet Allocation \cite{AnandkumarLDA},
and parsing models \cite{HsuKakadeLiang2012}.
The basic idea of this work is to express
the unknown model parameters as a tensor factorization
of the third-order moments of the model distribution, which
can be estimated from data.
The moments have a special symmetric structure
which permits the factorization to be computed efficiently using the robust
tensor power method \cite{AnandkumarGeHsu2012}.
%\cite{hsu09spectral}

% Twist
In mixture of linear regressions, using third-order moments does not directly
reveal the tensor structure of the problem, so we cannot directly
apply the above tensor factorization techniques.
Our solution is to employ low-rank linear regression \cite{NegahbanWainwright2009,Tomioka2011} on second
and third powers of the response to first recover the appropriate symmetric
tensors, from which we can then apply the tensor power method to retrieve the
final parameters.

% Approach: Derive moment functions based on linear regression on Y^2
% and Y^3 Algorithm: trace norm regularization to recover tensor
% + tensor factorization
%We propose an algorithm in \sectionref{sec:algo} that derives moment
%functions based on linear regression on the conditional moments of $y$,
%i.e. $\E[y^2 | X]$ and $\E[y^3 | X]$. We exploit the low rank properties
%of the second and third order moments to efficiently recover parameters,
%and then recover the individual regressors, $\beta_k$, using the tensor
%factorization method \cite{AnandkumarGeHsu2012}.

%These estimators feature polynomial
%sample and computational complexities while being statistically
%consistent, i.e. given sufficient data they are guaranteed to converge
%to the global optimum in an identifiable model, thus sidestepping the
%local optima issue altogether.

@@

% Show some identifiability results; require quadratic and cubic
% independence
%We prove that this approach is indeed statistically consistent with
%polynomial sample complexity.
% PSL: leave out for now (too detailed)
%Aside from
%standard identifiability results \citationneeded, we also show that
%recovery using moments places some independence requirements on the
%variables $X$.

@@

% Results
The result is a simple and efficient two-stage algorithm,
which we call Spectral Experts.
We prove that our algorithm yields consistent parameter estimates under certain
new identifiability conditions.  We also conduct an empirical evaluation
of our technique to understand its statistical properties (\sectionref{sec:evaluation}).
In particular, we find that in the low data regime,
Spectral Experts underperforms EM due to its weaker statistical efficiency,
but in the high data regime,
Spectral Experts outperforms EM.
We get the best results by using Spectral Experts as an initialization
to EM.

%We evaluate our algorithm on simulated linear and non-linear data to
%understand how well the algorithm scales with the number of components,
%dimensions and separation.

%To show how robust the algorithm is to model
%misspecification, we evaluate it's performance on \todo{a motion
%tracking dataset(?)}.

@@

%We compare our approach with expectation-maximization; our results,
%presented in \sectionref{sec:evaluation} shows that spectral methods do
%indeed provide an answer close enough to the global optima that using
%a local method from there recovers the true parameters. \todo{Describe
%some of our other high-level findings.}

%PSL: not for this paper
%\todo{What about talking about initialization? Generalizing and kmeans++}

@@

We use $x\tp{p}$ to represent the $p$-th order tensor
formed by taking the outer product of $x \in \Re^d$; i.e. $x\tp{p}_{i_1 i_2 \ldots
i_p} = x_{i_1} x_{i_2} \cdots x_{i_p}$. We will use
$\innerp{\cdot}{\cdot}$ to denote the generalized dot product between two $p$-th
order tensors: $\innerp{X}{Y} = \sum_{i_1, i_2, \ldots i_p} X_{i_1,
i_2, \ldots i_p} Y_{i_1, i_2, \ldots i_p}.$
For
$X \in (\Re^d)\tp{p}$ and $Y \in (\Re^d)\tp{(p+q)}$,% $p, q > 0$,
$\innerp{X}{Y} \in \Re^{d^q}$ is the projection of $X$ onto the first
$p$ modes of $Y$, i.e. $\innerp{X}{Y}_{i_1, \ldots, i_q}
= \sum_{j_1, \ldots, j_p} X_{j_1, \ldots, j_p} Y_{j_1, \ldots, j_p, i_1,
\ldots, i_q}$.
Let $[n] = \{ 1, \dots, n \}$ denote the first $n$ positive integers.
%Similar to how $x^T A$ is
%the projection of a vector $x$ onto the row space of a matrix $A$

%The expected moments of the parameters $\beta_1, \dots, \beta_k$ play an
%important role in our algorithm and we will use $M_p = \E[\beta\tp{p}]
%= \sum_{k=1}^{K} \pi_k \beta_k^{\otimes p}$ to denote them.
